\chapter{Programming}

This chapter describes the various ways and pieces the ZBox system is
programmed and customized.

\section{Objects}

\subsection{ObjectClass}

ZBox knows four kinds of Types, all derived from \emph{DataType}:
\begin{description}
\item[ObjectClass] { Type for ZBox Object Classes }
\item[Interface] { Type for ZBox Interfaces }
\item[Enumeration] { Type for ZBox Enumerations }
\item[CompoundObject] { Type for ZBox Object Classes }
\end{description}

\subsubsection{DataType}
This is the abstract base class for all ZBox Types. It provides the necessary
infrastructure to describe a Type.

\begin{CS}
public interface DataType 
{
    string Name;
    string Description;

    IList<Property> Properties;
    ICollection<Method> Methods;
    ICollection<MethodInvocation> MethodInvocations;

    Icon DefaultIcon;
    ICollection<InstanceConstraint> Constraints;
}
\end{CS}

\begin{description}
\item[Name] { The name of the \emph{DataType}. Note that this name has to be a
valid C\# name. A \emph{Constraint} protects this. }
\item[Description] { Each Type should have a description. This description is
used for documentation purposes. }
\item[Properties] { Each Type can have Properties (except \emph{Enumeration})}
\item[Methods] { Each Type can have Methods (except \emph{Enumeration})}
\item[MethodInvocations] { Methods on a \emph{DataType} are invoked by an
MethodInvokation. It does not matter on which level of type hierarchy the
invokation is defined. }
\item[DefaultIcon] { Each \emph{DataType} can have a default icon}
\item[Constraints] { A \emph{Constraint} checks the validity of an instance. If
any constraint throws an error nothing will be committed. }
\end{description}

\subsubsection{ObjectClass}
\emph{ObjectClass} is \emph{the} defining class for ZBox Objects.

\begin{CS}
public interface ObjectClass : DataType 
{
    ObjectClass BaseObjectClass;
    ICollection<Interface> ImplementsInterfaces;

    string TableName;

    bool IsAbstract;
    bool IsFrozenObject;
    bool IsSimpleObject;

    ViewModelDescriptor DefaultViewModelDescriptor;
    ControlKind RequestedKind;

    ICollection<AccessControl> AccessControlList;
}
\end{CS}

\begin{description}
\item[IsAbstract]
\item[IsSimpleObject] The \emph{ObjectClass} is marked as a simple, non complex
object. Set it to true for all classes that have no inheritance and less
enought and simple properties to edit them in grids and/or lists. This will
prevent all list view models from opening such an object directly. Also the \emph{ObjectBrowser} will not provide any interface.
\end{description}

\input{Programming/Relations}

\subsection{Properties}

\begin{CS}
public interface Property 
{
    string Name;
	...
}
\end{CS}

\begin{description}
\item[Editable in Grid] This is an override for the inline editing behaviour for
grid and list controls. The default behaviour is to evaluate if the referenced
\emph{ObjectClass} is a simple object or not.
\end{description}

\subsection{Additional Metadata}

The object model is intended to be very rich and provide the various
subsystems with meta data directly from the \emph{ObjectClass}.

This section describes the various pieces of this meta data.

\subsubsection{New related objects}

A \emph{CreateRelatedUseCase} describes the use case of creating a new
object related to the "current" instance. One such use case would be
e.\,g.~"create a new \emph{Relation} from the current ObjectClass."

Such use cases are described with \emph{CreateRelatedUseCase} objects:

\begin{CS}
interface CreateRelatedUseCase
{
	string Label;
	Method Action;
	Relation AffectedRelation; // optional
}
\end{CS}

The \emph{Action} will be called when the user requests an execution of
this use case. This method doesn't take any parameters and returns the
newly created object. The infrastructure on the client will cause the
returned object to be displayed to the user. The business logic should
already have filled out the property values according to the use case.
The name of the method should start with "Create".

If the optional \emph{AffectedRelation} is specified, one of its ends it
must match the \emph{ObjectClass} of the \emph{Method}. This relation
can then be used to identify controls in the UI where the action can be
placed.

\section{Modules}

\section{Compound Objects}
Lets say there is a \emph{PhoneNumber} Compound Object.
\begin{CS}
class PhoneNumber
{
	string CountryCode;
	string AreaCode;
	string Number;
	string Extension;
}
\end{CS}

A Person has two phone numbers:

\begin{CS}
class Person
{
	string Name;
	...
	PhoneNumber  Tel;
	PhoneNumber? Fax;
}
\end{CS}
\emph{Tel} is not nullabe, \emph{Fax} is nullable.

\subsection{Accessing Compound Objects}

\begin{itemize}
	\item{If a compound object property is not nullable then the content of the property is always a valid reference.}
	\item{If a compound object property is  nullable then the content of the property may be null.}
	\item{When a compound object property is set the given compound object will be copied.}
\end{itemize}

\begin{CS}
Person p;
string number;

number = p.Tel.Number;
number = p.Fax.Number; // throws NullReferenceException if Fax is null

p.Tel.Number = "12345678";
p.Fax.Number = "12345678"; // throws NullReferenceException if Fax is null

PhoneNumber n;

n = p.Tel; // returnes a reference of the compound object
n.Number = "87654321"; // changes p.Tel.Number
n = p.Fax; // may be null
n.Number = "87654321"; // changes p.Fax.Number or throws NullReferenceException if Fax is null

p.Fax = p.Tel; // creates a copy of p.Tel
p.Fax.Number = "87654321"; // changes p.Fax.Number but does not change p.Tel.Number

p.Tel = null; // throws a ArgumentNullException
p.Fax = null; // sets Fax to null

n = ctx.CreateStruct<PhoneNumber>(); // creates a new PhoneNumber Struct;
n.Number = "12345678";
p.Tel = n; // creates a copy of n
p.Tel.Number = "18273645"; // changes p.Tel.Number but does not change n.Number
n.Number = "87654321"; // changes n.Number but does not change p.Tel.Number
\end{CS}

\section{Enhancing Kistl's inner workings}

\subsection{Database Providers}

\input{Programming/GraphicalUserInterface}

\section{Core Kistl Development Environment}

\subsection{Preparing a clean local build}

First, it is necessary to have a clean build environment. Use
\emph{subst} to create a drive \emph{P:} where your checkout
resides in a directory called \emph{Kistl}.

The \emph{!FullReset.cmd} will bring the database and the bootstrapping
code up to the current \emph{Database.xml}'s content.

Now the environment is ready for programming.

\subsection{Merging local and remote changes}

When the subversion repository has changed the \emph{Database.xml} while
local changes were made to the schema, it is necessary to merge them
before comitting.

After fetching and merging the update from the subversion repository,
the local \emph{Database.xml} has changes which are not yet in the
database. Running \emph{!DeployAll.cmd} updadates the SQL-schema and
produces a new set of generated assemblies in the \emph{CodeGenPath}.
After testing that the merge was successful, use \emph{GetCodeGen.cmd}
to update the working directory with the newly generated bootstrapping
code.

Now the working directory is ready for check in.



